---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Q1

```{r,1}
Xij <- scan(text= ## Data Taylor & Ashe (1983)
"357848 0766940 0610542 0482940 527326 574398 146342 139950 227229 067948
352118 0884021 0933894 1183289 445745 320996 527804 266172 425046
290507 1001799 0926219 1016654 750816 146923 495992 280405
310608 1108250 0776189 1562400 272482 352053 206286
443160 0693190 0991983 0769488 504851 470639
396132 0937085 0847498 0805037 705960
440832 0847631 1131398 1063269
359480 1061648 1443370
376686 0986608
344014")
n <- length(Xij); TT <- 10
i <- rep(1:TT, TT:1); j <- sequence(TT:1)
i <- as.factor(i); j <- as.factor(j)
Orig.CL <- glm(Xij~i+j, quasipoisson)
coefs <- exp(coef(Orig.CL))
alpha <- c(1, coefs[2:TT]) * coefs[1]
beta <- c(1, coefs[(TT+1):(2*TT-1)])
names(alpha) <- paste0("row",1:10); round(alpha)
names(beta) <- paste0("col",1:10); round(beta, 4)
#-----------------------------------------------------------------------
#-----------------------------------------------------------------------
(tapply(Xij,i,sum) - tapply(fitted(Orig.CL),i,sum))/tapply(Xij,i,sum)
(tapply(Xij,j,sum) - tapply(fitted(Orig.CL),j,sum))/tapply(Xij,j,sum)
#-----------------------------------------------------------------------
#-----------------------------------------------------------------------
```

## Q2

row(Orig.fits):Returns a matrix of integers indicating their row number in the Orig.fits

col(Orig.fits):Returns a matrix of integers indicating their col number in the Orig.fits

future: the logical matirx which True below the diagonal and FLASE above the diagonal
```{r,2}
Orig.fits <- alpha %o% beta; round(Orig.fits) ## display fitted values
future <- row(Orig.fits) + col(Orig.fits) - 1 > TT
Orig.reserve <- sum(Orig.fits[future]) ## 18680856
print('------result------')
row(Orig.fits)
col(Orig.fits)
future
```

##Q3:
Here are 121 negative pseudo-observations appeared in total.
```{r,3}
Prs.resid <- (Xij - fitted(Orig.CL)) / sqrt(fitted(Orig.CL))
p <- 2*TT-1; phi.P <- sum(Prs.resid^2)/(n-p)
Adj.Prs.resid <- Prs.resid * sqrt(n/(n-p))
#-----------------------------------------------------------------------
#-----------------------------------------------------------------------
set.seed(1) ## fill in your favorite 6-digit number
#-----------------------------------------------------------------------
#-----------------------------------------------------------------------
nBoot <- 1000; payments <- reserves <- n.neg <- numeric(nBoot)
number = 0
for (boots in 1:nBoot){
  Ps.Xij <- sample(Adj.Prs.resid, n, replace=TRUE) ## 1
  Ps.Xij <- Ps.Xij * sqrt(fitted(Orig.CL)) + fitted(Orig.CL) ## 2
  number.neg <- sum(Ps.Xij<0)
  for(a in Ps.Xij){if (a<0){number = number+1}}
  Ps.Xij <- pmax(Ps.Xij, 0) ## Set obs < 0 to 0
  Ps.CL <- glm(Ps.Xij~i+j, family=quasipoisson) ## 5
  coefs <- exp(as.numeric(coef(Ps.CL)))
  Ps.alpha <- c(1, coefs[2:TT]) * coefs[1]
  Ps.beta <- c(1, coefs[(TT+1):(2*TT-1)])
  Ps.fits <- Ps.alpha %o% Ps.beta
  Ps.reserve <- sum(Ps.fits[future])
  h <- length(Ps.fits[future]) ## 11
  Ps.payments <- phi.P * rpois(h, Ps.fits[future]/phi.P)
  Ps.totpayment <- sum(Ps.payments)
  reserves[boots] <- Ps.reserve ## 14
  payments[boots] <- Ps.totpayment
  n.neg[boots] <- number.neg
}
number
```

##Q4
```{r,4}
PE.bs <- sqrt(phi.P*Orig.reserve + var(reserves)) ## ~ 3 million
var(reserves) / (phi.P*Orig.reserve) ## ~ 8
payments <- payments/1e6
mean(payments) ## sample mean: ~ 19
sd(payments) ## sample s.d.: ~ 3
pp <- payments-mean(payments)
mean(pp^3)/sd(payments)^3 ## sample skewness: ~ 0.4, varies a lot
mean(pp^4)/mean(pp^2)^2 - 3 ## sample kurtosis: ~ 0.4, varies a lot
hist(payments, breaks=21, prob=T)
lines(density(payments), col="blue")
curve(dnorm(x, mean(payments), sd(payments)), add=T, col="red")
quantile(payments, c(0.5,0.75,0.9,0.95,0.99))
```

##Q5:

```{r,5}
#(a)
Orig.gam <- glm(Xij~i+j, fam=Gamma(link=log))
#(b)
coefs <- exp(coef(Orig.gam))
alpha = c(1, coefs[2:TT]) * coefs[1]
beta <- c(1, coefs[(TT+1):(2*TT-1)])
names(alpha) <- paste0("row",1:10); round(alpha)
names(beta) <- paste0("col",1:10); round(beta, 4)
Orig.fits <- alpha %o% beta; round(Orig.fits) ## display fitted values
future <- row(Orig.fits) + col(Orig.fits) - 1 > TT
Orig.reserve <- sum(Orig.fits[future]) #18085805
Orig.reserve
#(c)+(d)
Prs.resid <- (Xij - fitted(Orig.CL)) / sqrt(fitted(Orig.CL))
p <- 2*TT-1; phi.P <- sum(Prs.resid^2)/(n-p)
Adj.Prs.resid <- Prs.resid * sqrt(n/(n-p))
set.seed(1) ## fill in your favorite 6-digit number
nBoot <- 10; payments.gam <- reserves.gam <- n.neg <- numeric(nBoot)
for (boots in 1:nBoot){
  Ps.Xij <- sample(Adj.Prs.resid, n, replace=TRUE) ## 1
  Ps.Xij <- Ps.Xij * sqrt(fitted(Orig.gam))+ fitted(Orig.gam) ## 2
  number.neg <- sum(Ps.Xij<0)
  Ps.Xij <- pmax(Ps.Xij, 0.1) ## Set obs < 0 to 0
  print(Ps.Xij)
  Ps.CL <- glm(Ps.Xij~i+j, family=Gamma(link=log)) ## 5
  coefs <- exp(as.numeric(coef(Ps.CL)))
  Ps.alpha <- c(1, coefs[2:TT]) * coefs[1]
  Ps.beta <- c(1, coefs[(TT+1):(2*TT-1)])
  Ps.fits <- Ps.alpha %o% Ps.beta
  Ps.reserve <- sum(Ps.fits[future])
  h <- length(Ps.fits[future]) ## 11
  Ps.payments <- phi.P * rpois(h, Ps.fits[future]/phi.P)
  Ps.totpayment <- sum(Ps.payments)
  reserves.gam[boots] <- Ps.reserve ## 14
  payments.gam[boots] <- Ps.totpayment
  n.neg[boots] <- number.neg
}
#(e)
PE.bs <- sqrt(phi.P*Orig.reserve + var(reserves.gam)) ## ~ 3 million
var(reserves.gam) / (phi.P*Orig.reserve) ## ~ 8
payments.gam <- payments.gam/1e6
mean(payments.gam) ## sample mean: ~ 19
sd(payments.gam) ## sample s.d.: ~ 3
pp <- payments.gam-mean(payments.gam)
mean(pp^3)/sd(payments.gam)^3 ## sample skewness: ~ 0.4, varies a lot
mean(pp^4)/mean(pp^2)^2 - 3 ## sample kurtosis: ~ 0.4, varies a lot
quantile(payments.gam, c(0.5,0.75,0.9,0.95,0.99))
```

## Q6

From the result, the parameters alpha and beta extracted from Orig.gam satisfy the DM equations. 

```{r,6,warning=FALSE}
alpha
beta
w = coefs
for (iter in 1:10){
  alpha <- sqrt(tapply(w*Xij/beta[j],i,sum)/
  tapply(w*beta[j],i,sum))
  beta <- sqrt(tapply(w*Xij/alpha[i],j,sum)/
  tapply(w*alpha[i],j,sum))}
alpha
beta
```

## Q7

(a)

the output look as my expect, the output o those are the same in the first item, and add zeros for future years in Xij.

(b)

We can see that the coef and deviance are the same in these two model CL and Orig.CL
```{r,7}
Xij.1 <- as.vector(t(xtabs(Xij~i+j))) ## stored row-wise as usual
ii <- rep(1:TT, each=TT); jj <- rep(1:TT, TT); future <- ii+jj-1 > TT
ii <- as.factor(ii); jj <- as.factor(jj)
CL <- glm(Xij.1~ii+jj, fam=quasipoisson, wei=as.numeric(!future))
Xij[i==TT]
Xij.1[ii==TT]
coef(CL)
coef(Orig.CL)
CL$deviance
Orig.CL$deviance
```

##Q8£º

line 1 is equivalent to the mean deviance estimate 

line 2 is equivalent to the Pearson estimate

```{r,8}
n <- sum(Xij.1>0) ## number of observations
p <- 2*TT-1 ## number of estimated parameters
phi <- CL$deviance/CL$df.residual; phi ## 1
1/(n-p)*sum((!future)*(Xij.1 - fitted(CL))^2/fitted(CL)) ## 2
sum(resid(CL,type="devi")^2)/(n-p) ## 3
sum(resid(CL,type="pear")^2)/(n-p) ## 4
2/(n-p)*sum((Xij.1*log(Xij.1/fitted(CL)) - (Xij.1-fitted(CL)))[!future]) ## 5
summary(CL)$dispersion ## 6
Cov.beta <- vcov(CL)
X <- model.matrix(CL)
Cov.eta <- X %*% Cov.beta %*% t(X)
mu.hat <- fitted(CL)*future ## predictions for future cells
MSE <- phi * sum(mu.hat) + t(mu.hat) %*% Cov.eta %*% mu.hat ## equation (1)
sqrt(MSE) ## 2946484 = Root MSE = prediction error
sum(mu.hat)
```

##Q9:

{r,9}
mu.hat
for (r in 2:TT){
mu.r <- mu.hat ## replace the elements of mu.hat not having rownr==r by 0
res <- .... ## reserve estimate for year of origin r
MSE <- .... ## MSE for year of origin r, see above
cat("Year =", r, "\t reserve =", round(res), "\t pred.error/reserve =",
round(100*sqrt(MSE)/res), "%\n")
}


##Q10:

The scaled deviance of exposure model is 7.1 compared with scaled deviance of chain ladder model is 8.83, and in the freedom , there has no difference, so chain ladder model comes out best
```{r,10}
rm(list=ls())
Xij <- scan(text="
156 37 6 5 3 2 1 0
154 42 8 5 6 3 0
178 63 14 5 3 1
198 56 13 11 2
206 49 9 5
250 85 28
252 44
221")
TT <- 8; i <- rep(1:TT, TT:1); j <- sequence(TT:1)
fi <- as.factor(i); fj <- as.factor(j)
ee <- c(28950,29754,22315,39442,38423,50268,44762,43541)
Expo <- ee[i]
CL <- glm(Xij~fi+fj, quasipoisson) ## chain ladder model
EE <- glm(Xij~offset(log(Expo))+fj, quasipoisson) ## exposure model
phi <- CL$deviance / CL$df.residual ## mean-deviance estimate for phi
Delta.Dev.Sc <- 2*log(CL$deviance) ## difference of scaled deviances for CL and EE
Delta.df <- CL$df.null-EE$df.null  ## difference of degrees of freedom for CL and EE
reject <- ifelse(Delta.Dev.Sc>phi**2, TRUE, FALSE) ## TRUE if Delta.Dev.Sc > the chi^2 critical value
cat("The exposure model", ifelse(reject, "is", "is not"), "rejected",
"since the scaled deviance gained by CL is\n",
round(Delta.Dev.Sc,1), "with", Delta.df, "extra parameters.\n") 
2*log(EE$deviance)
```

## Q11

```{r,11}
xtabs(round(100*(fitted(CL) - fitted(EE))/fitted(CL))~i+j)
round(coef(CL),2); round(coef(EE),2)
CL.off <- glm(Xij~offset(log(Expo))+fj+fi, quasipoisson)
summary(CL.off)
exp(coef(CL))
exp(coef(CL.off))
summary(CL.off)
```

##Q12:

(a)

The adjusted exposure model yields significant improvement over the ordinary exposure model, since there is some change in the deviance.

(b)

chain ladder model do not yields significant improvement over the adjusted exposure model, since there is little change (0.85293<1) in the deviance.

(c)
From the result (a) and (b), the best model is EE.adj

```{r,12}
i.is.3 <- as.numeric(i==3)
EE.adj <- glm(Xij~offset(log(Expo))+i.is.3+fj, quasipoisson)
anova(EE, EE.adj)
anova(EE.adj, CL)

```


##Q13:

Similar to the above analysis, we can get similar results.Compared with the result of Q12, The Deviance of ANOVA (EE1, EE1.adj) reduced from 47.874 to 2.9203, which also shows that EE.adj has a less prominent advantage, but it is still the best model
```{r,13}
ee1 <- c(28950,29754,32315,39442,38423,50268,44762,43541)
Expo1 <- ee1[i]
CL1 <- glm(Xij~fi+fj, quasipoisson) ## chain ladder model
EE1 <- glm(Xij~offset(log(Expo1))+fj, quasipoisson) ## exposure model
EE1.adj <- glm(Xij~offset(log(Expo1))+i.is.3+fj, quasipoisson)
anova(EE1, EE1.adj)
anova(EE1.adj, CL1)
```

##Q14:

```{r,14}
rm(list=ls()) ## clear workspace
Xij <- scan(text="
156 37 6 5 3 2 1 0
154 42 8 5 6 3 0
178 63 14 5 3 1
198 56 13 11 2
206 49 9 5
250 85 28
252 44
221")
TT <- 8; i <- rep(1:TT, TT:1); j <- sequence(TT:1)
fi <- as.factor(i); fj <- as.factor(j)
ee <- c(28950,29754,31141,32443,34700,36268,37032,36637)
Expo <- ee[i]
CL <- glm(Xij~fi+fj, quasipoisson)
cc <- exp(coef(CL))
alpha <- cc[1]*c(1,cc[2:TT]); beta <- c(1,cc[(TT+1):(2*TT-1)])
alpha <- alpha*sum(beta); beta <- beta/sum(beta)
M <- ee * alpha[1] / ee[1]
```

## Q15:

From the result, CL.retro is equal to sum(Xij), but BF.retro is not

Reason:

```{r,15}
####(a)
CL.fits <- alpha %o% beta; round(CL.fits, 2)
BF.fits <- M %o% beta; round(BF.fits, 2)
future <-  row(CL.fits) + col(CL.fits) - 1>TT ## see code preceding Q2
CL.reserve <-  sum(CL.fits[future]); CL.reserve
BF.reserve <-  sum(BF.fits[future]); BF.reserve
####(b)
CL.retro <- sum(CL.fits[future-1])
print('-----CL.retro-----')
CL.retro
BF.retro <- sum(BF.fits[future-1])
print('-----BF.retro-----')
BF.retro
####(c)
print('------XIJ SUM------')
sum(Xij)
```



